Order Preserving Encryption Schemes: \\ Preventing the Leakage of Distribution\\ 6.857 Final Paper \\ Hrishikesh Joshi, Max Kolysh, Mari Miyachi, John Wang

### 2013-05-06 paper.tex
Proving equivalence of the schemes.

\begin{theorem}
  DCS is IND-CPDA secure 
\end{theorem}
\begin{proof}
We shall show that encrypting $v = \{V_1, V_2, \ldots, V_{\kappa} \}$ and $w = \{W_1, W_2, \ldots, W_{\kappa}\}$ is information-theoretically the same, where $V_i, W_i$ are sequences of size $\lambda$.

@@

  RTOPE can be proven to be IND-CPDA, just like the DCS scheme. In fact, one can show the equivalence of RTOPE and DCS in terms of the OPE encodings that both provide.

@@

  We shall proceed to show that the two schemes are identical as the number of repetitions increases to infinity.

  \begin{theorem}
    DCS and ROPTE are equivalent when the number of repetitions of each distinct repeated value tends to infinity.
  \end{theorem}
  \begin{proof}
    We shall prove this theorem by examining a sequence of moves when looking at the insertion of the $n$th repeated value. If we can show that the probability of traversing the tree in a particular sequence $\{m_i\}_{i=0}^k$ is the same in both DCS and RTOPE, then the two schemes can be shown to be equivalent. We shall show that as $n \to \infty$, the probability of obtaining a particular sequence converges to the same value in both DCS and RTOPE.

  Denote $x$ as the $n$th repeated value. In DCS, to move right so that $m_i = 1$, one must have $p_i < x$. To move left so that $m_i = 0$, one must have $p_i > x$. The probability of choosing a value $x$ to move left is just $p_i$ (probability to pick a value below $p_i$) and moving right is $1 - p_i$ (probability to pick a value above $p_i$). Thus, the probability of making move $m_i$ is $(1 - m_i) p_i + m_i (1 - p_i)$.

\end{proof}

>>>>

\begin{theorem}
  DCS is IND-CPDA secure.
\end{theorem}
\begin{proof}
We shall show that encrypting $v = \{V_1, V_2, \ldots, V_{\kappa} \}$ and $w = \{W_1, W_2, \ldots, W_{\kappa}\}$ is information-theoretically the same, where $V_i, W_i$ are sequences of size $\lambda$.

@@

@@

  We shall proceed to show that the two schemes are actually identical. In order to do this, we need to first prove a lemma.
  \begin{lemma}
    Let $\{ x_i \}_{i=1}^n$ be a sequence of $n$ independent uniform random variables chosen from the interval $[0,1)$. Let $x_{n+1}$ be another uniform random variable on the interval $[0,1)$. The probability that $x_{n+1}$ is the $j$th smallest item is $1/(n+1)$.
  \end{lemma}
  \begin{proof}
    For the proof, we rely on the fact that the $n$ random variables are independent. We note that the $n+1$ random variable can be thought of as the next random variable in the sequence. Moreover, since all of the random variables are identically drawn from the same uniform distribution, we see that we are really drawing from a sequence of $n+1$ uniform random variables which can be chosen at the same time.

    It is clear that the probability of any $x_i$ being the $j$th smallest item is the same, because all of the random variables are drawn independently. This means that $Pr[x_{n+1} = X_{j}]$ where $X_j$ is the $j$th smallest item can be given by the expression:
    \begin{eqnarray}
      Pr[x_{n+1} = X_{j}] = Pr[x_{n} = X_{j} = \ldots = Pr[x_1 = X{j}]
    \end{eqnarray}

    Since exactly one $x_{i}$ is the $j$th smallest element, we see that $\sum_{i} Pr[x_i = X_j] = 1$. This implies that $Pr[x_i = X_j] = 1/(n+1)$, which further implies that $Pr[x_{n+1} = X_j] = 1/(n+1)$. This completes the proof.
  \end{proof}

  This lemma allows us to show that DCS and RTOPE have identical move distributions, and thus are actually equivalent.

  \begin{theorem}
    DCS and ROPTE are equivalent.
  \end{theorem}
  \begin{proof}
    We shall prove this theorem by examining a sequence of moves when looking at the insertion of a repeated value. If we can show that the probability of traversing the tree in a particular sequence $\{m_i\}_{i=1}^k$ is the same in both DCS and RTOPE, then the two schemes can be shown to be equivalent. We shall show that as $n \to \infty$, the probability of obtaining a particular sequence converges to the same value in both DCS and RTOPE.

  Denote $x$ as the repeated value to insert. In DCS, to move right so that $m_i = 1$, one must have $p_i < x$. To move left so that $m_i = 0$, one must have $p_i > x$. Although there exists a sequence of moves from the root to a leaf of the tree which includes every node in the tree, we need only examine the nodes which have a plaintext value equal to $x$. This is because the traversal of DCS and RTOPE is the same for nodes which are not repeated (they share the normal binary tree traversal rule). Thus, we know that the moves in these cases will be equivalent and that the probability of making the moves on non-repeated nodes will be equal.

  Thus, we can reduce our problem to considering a sequence of moves $\{m_i\}_{i=0}^k$ on the plaintext values which are equal to $x$. If we can show that the probability of making a particular sequence of moves on the repeated values is the same for DCS and RTOPE, then we will be done. Making any sequence of moves on repeated values in RTOPE will have probability $2^{-k}$ where $k$ is the number of repeated values in the sequence. This is because one moves left or right with probability $\frac{1}{2}$.

  Now, we shall examine the sequence of moves in RTOPE. Recall that we only need to consider a sequence of moves $\{m_i\}_{i=0}^k$ which is performed on a single repeated plaintext value $x$. This means that in RTOPE, the plaintext values will be distinguished by the random bit $r_i$ so that $p_{i} = x + r_i$. Therefore, we need only compare $r_i$ to the value $x + r_{k+1}$ that we are querying.

  However, notice that $\{r_i\}_{i=1}^k$ is a sequence of $k$ independent uniform random variables chosen from the interval $[0,1)$. Moreover, $r_{k+1}$ is chosen from the same interval and is also an independently drawn uniform random variable. This means, that the probability that $r_{k+1}$ is the $i$th smallest element is $1/(k+1)$. Therefore, it is equally probable that the next move $x$ ends up going through any sequence of moves down the tree. This follows because each rank corresponds to a different path through the tree, and a different sequence of moves. Thus, each sequence $\{m_i\}_{i=1}^k$ is equally probable. Since there are $2^k$ different possible sequences, the probability of going down any such sequence is $2^{-k}$, which is the same probability as in RTOPE.

    Thus, we have shown that the probability of inserting a new repeated element of $x$ in any direction gives the same probability distribution in both DCS and RTOPE. Since the insertion of a non-repeated element will give rise to the same tree, if the tree was the same before, we see that the probability of obtaining a particular OPE encoding is the same in both DCS and RTOPE. Thus, we can take the bijective mapping $f(x) = x$ to use in the definition of equivalence. It is then clear that $Pr[Encode_{DCS}(c_i) = \omega] = Pr[Encode_{RTOPE}(c_i) = \omega]$ for all $\omega \in \mathrm{Z}_x$. This proves equivalence.
\end{proof}

  RTOPE can be proven to be IND-CPDA, just like the DCS scheme. In fact, now that we have shown that DCS is IND-CPDA secure and that RTOPE and DCS are equivalent, proving the security of RTOPE is trivial. We have already seen that the equivalence of two schemes imply that one is secure if and only if the other is secure. Thus, we can prove the theorem:

  \begin{theorem}
    RTOPE is IND-CPDA secure.
  \end{theorem}
  \begin{proof}
    We have shown that RTOPE and DCS are equivalent. We have also shown that if $x,y$ are equivalent, then $x$ is IND-CPDA secure if and only if $y$ is IND-CPDA secure. Since we have shown that DCS is IND-CPDA secure, we see that RTOPE is also IND-CPDA secure.
  \end{proof}
